---
title: "WUP_04"
author: "Julianne Mercado"
date: "fall 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
mydata<-read.csv(file="GSScsv.csv")
```
```{r}
boxplot(WORDSUM~DEGREE,data=mydata)
```

The middle box and the far right box can go completely under the box to the left labeled "graduate." The means from the two sets are different from each other, therefore if you run any test with the two boxes (the one in the middle and the one far right), you can suspect that their mean values for thew WORDSUM test are a lower value then the boxes "Bachelor" "Graduate" and "Junior College". 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```

The individuals with a janitor college degree, a bachelors degree, and a graduate school degree have higher mean result than the high school and junior high school categories. As you can see in the plot, the individuals with degrees have a value of 9 or higher, while the LT High School students and the high school students have a mean value of 8 or less. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```
The P value seems very small. The P value represents the chances of getting this result or something farther when the null hypothesis is actually true. I do not believe that, therefore i reject the null hypothesis. 

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1,conf.level=.95),las=2)
```

This shows that high school and graduate are different from each other and LT high school and graduate are different from each other. We can see that LT high school- graduate and high school-graduate do not hit the "g" line. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

  In this boxplot we discussed how individuals with a graduate, bachelor degree, and a junior college degree scored better than individuals in the LT high school category and the high school category. We established that the people who attended college had a mean value of 9 or greater, while the LT high school and the high school students had a mean value of 8 or less. We found that our p value was a very small value of 0.000236, so rejected the null hypothesis. Lastly we used the "TukeyHSD" model and were able to elaborate further and see that at the 95% confidence level LT high school- graduate and high school- graduate categories did not go over the "0". In conclusion these two categories referenced are different. 